 
\section{Another Multiplayer Experiment}\label{VDexperiment}

This activity needs to be played as a class. All players need to be able to respond without being able to see the responses of others. Answers may be revealed before moving on to the next question.

Without sharing your answers with others, select your answer to the following questions. Try to be as honest about your answer as possible.  Make sure you have a reason for each answer.

\begin{enumerate}
\item The lights go out in the neighborhood. Someone needs to call the power company. If someone calls, everyone's lights go on.

\begin{enumerate}
\item Call
\item Don't call
\end{enumerate}

\item The same as in (1), but now you have to wait on hold for 5 minutes.

\begin{enumerate}
\item Call
\item Don't call
\end{enumerate}

\item The same as in (1), but now you have to wait on hold for 30 minutes.

\begin{enumerate}
\item Call
\item Don't call
\end{enumerate}

\item The same as in (1), but now you have to pay a \$.50 service fee.

\begin{enumerate}
\item Call
\item Don't call
\end{enumerate}

\item The same as in (1), but now you have to pay a \$2.50 service fee.

\begin{enumerate}
\item Call
\item Don't call
\end{enumerate}

\item The phone lines go down in your small mountain community. You have to hike 3 miles in the snow to notify the power company.

\begin{enumerate}
\item Hike to notify the phone company
\item Stay home and let someone else do the hiking
\end{enumerate}

\item Everyone in your class cheats on an exam. The professor says if someone confesses they receive an F. If no one confesses, everyone fails.

\begin{enumerate}
\item Confess
\item Don't confess
\end{enumerate}

\item Evil Dr. No captures the class and puts you in all in a shark tank separated so you can't communicate. If one person volunteers to be eaten then the rest go free. If no one volunteers after 10 minutes all get eaten by sharks.

\begin{enumerate}
\item Volunteer
\item Don't volunteer
\end{enumerate}

\item Evil Dr. No captures your family and puts you in all in a shark tank separated so you can't communicate. If one person volunteers to be eaten then the rest go free. If no one volunteers after 10 minutes all get eaten by sharks.

\begin{enumerate}
\item Volunteer
\item Don't volunteer
\end{enumerate}

\item For any ``Big Brother'' fans. Choose to eat all your favorite foods for a week or nasty ``slop'' for a week. If at least three people say slop, everyone gets what they asked for. Otherwise everyone is on slop.

\begin{enumerate}
\item Favorite foods
\item Slop
\end{enumerate}

\item OK, now let's get serious about this. Answer 5 points or 1 point. If at least one person says 1 point, everyone gets the number of points they chose. Otherwise, everyone gets 0 points.

\begin{enumerate}
\item 5 points
\item 1 point
\end{enumerate}

\item Answer 20 points or 1 point. If at least one person says 1 point, everyone gets the number of points they chose. Otherwise, everyone gets 0 points.

\begin{enumerate}
\item 20 points
\item 1 point
\end{enumerate}

\item Answer 6 points or 5 points. If at least one person says 5 points, everyone gets the number of points they chose. Otherwise, everyone gets 0 points.

\begin{enumerate}
\item 6 points
\item 5 points
\end{enumerate}

\item Answer 20 points or 1 point. If at least \emph{five}  people say 1 point, everyone gets the number of points they chose. Otherwise, everyone gets 0 points.

\begin{enumerate}
\item 20 points
\item 1 point
\end{enumerate}
\end{enumerate}
-----------------------------------------------------


\begin{xca} After answering the questions, were you likely to volunteer or unlikely to volunteer? For example, were you likely to be the one to call the power company or get eaten by sharks, or were you generally hoping someone else would do it? If it depended on the situation, explain under what circumstances you were likely to volunteer.
\end{xca}

\begin{xca}
After sharing your answers as a class, did each situation have a volunteer? In other words, was there always someone willing to call the power company or take fewer points? If there was a question with no volunteer, can you suggest why?
\end{xca} 

\begin{xca}
For which questions was it unlikely that there would be very many volunteers? Did you take that into consideration when deciding if you were going to volunteer?
\end{xca}

\begin{xca}
What are some reasons for volunteering? What are some reasons not to volunteer?
\end{xca}